---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "65%"
)
```

# gbatr

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

**gbatr** is an interface to [Geosupport Dektop Edition](https://www1.nyc.gov/site/planning/data-maps/open-data/dwn-gde-home.page), a batch geocoder for New York City created by the NYC Department of City Planning. It enables very fast local geocoding of NYC addresses. This is a fork of the [rGBAT16AB pacakge](https://github.com/gmculp/rGBAT16AB) developed by GM Culp.

*NOTE: This is very much a work-in-progress and is very likely to change. If you would like to help with development, please send me a message or open an issue!*

## Example

To use gbatr to geocode addresses, first prepare a data frame that contains a column of street addresses and a column with either zip codes or boroughs. Note, this only works for addresses in New York City! The zip or borough column must all zip codes or all boroughs, not a mix of both.

Then, call the `gbat()` function and specify the input data frame, the name of the the address column, the name of the zip or borough column, whether you are using zip codes or boroughs, and which geocoded fields you want to get back from the GBAT geocoder. 

```{r}
library(gbatr)

za <- tibble::tribble(
  ~name,                 ~address,              ~borough,
  "Roberta's",           "261 Moore Street",    "Brooklyn",
  "L'Industrie",         "254 S 2nd Stret",     "Brooklyn",
  "Emmy Squared",        "364 Grand Street",    "Brooklyn",
  "Di Fara",             "1424 Avenue J",       "Brooklyn",
  "L&B Spumoni Gardens", "2725 86th Street",    "Brooklyn",
  "Totonno's",           "1524 Neptune Avenue", "Brooklyn"
  )

za_geo <- gbat(
  za,
  address = "address",
  zip_boro = "borough",
  zip_boro_type = "boro",
  geo_colnames = "lot_centroid_latlon"
  )

za_geo
```

## Why gbatr?

Geocoding is the process of converting places or addresses to geographic points, often latitudes and longitudes or x and y coordinates. It is a common process for geographers and data scientists because most often data sets contain street addresses of points of interest. In order to make maps and perform spatial analysis, these addresses must be first converted to geographic coordinates.

There are many tools to turn addresses into points. Some R packages that have geocoding functionality include [**ggmap**](https://github.com/dkahle/ggmap), [**censusxy**](https://slu-opengis.github.io/censusxy/index.html), and [**tidygeocoder**](https://jessecambon.github.io/tidygeocoder/). These are all good options, but one major drawback of all of them is that they use web APIs to perform the geocoding. This means that you send your addresses (either one by one or in a batch) to someone's webserver and you get back the geocoded points. Two consequences of this are 1) that large requests may be quite slow and 2) you may not be able to ensure the privacy of the addresses you send out (if they are sensitive). A third concern, depending on which geocoding service you use, is cost. In particular, Google has recently changed the cost structure of their [Geocoding API service](https://developers.google.com/maps/documentation/geocoding/usage-and-billing).

Given these challenges, it is fortunate that a fast, free batch geocoder that can be run locally exists. It is called [**Geosupport Desktop Edition**](https://www1.nyc.gov/site/planning/data-maps/open-data/dwn-gde-home.page) and was created by the [New York City Department of City Planning](https://www1.nyc.gov/site/planning/index.page). The major downside of this geocoder is that it only works for New York City address.

City Planning makes the geocoder available as a Windows and Linux desktop application (there are server version as well), in which you specify an Excel workbook (or other file type) of address, set some parameters, and the geocoder adds the geocoded information to a new sheet in the workbook. One issue with this method of using Geosupport is that you can't script or automate this process. If you are trying to develop a data pipeline in which one step is geocoding addresses, you would have to export your data, manually perform the geocoding, import the data, and continue your processing. So, if geocoding is a common task, gbatr can greatly simplify the task by using Geosupport directly from R with a data frame input and output.

## Installation

Before you install gbatr, you must have Geosupport Desktop Edition installed. Geosupport is a free geocoding application developed by the NYC Department of City Planning. It is available for Windows and Linux and can be downloaded from the [City Planning website](https://www1.nyc.gov/site/planning/data-maps/open-data/dwn-gde-home.page). For now, gbatr has only been tested using the Windows 64-bit version of Geosupport, so let me know if you're having troubling installing on Linux or 32-bit Windows.

When installing Geosupport Desktop, it is strongly recommended to use the default installation path (`C:/Program Files/Geosupport Desktop Edition`). If you install it in this location, building gbatr from source should just work. If you can't install it in this location, you may have to do some additional work to build gbatr.

Once you have Geosupport Desktop Edition installed, the easiest way to install gbatr is by using the [remotes package](https://remotes.r-lib.org/). In order to install using remotes, you will likely need to set an [environment variable to allow warnings during installation](https://github.com/r-lib/remotes#environment-variables) before you try to install:

```{r eval=FALSE}
Sys.setenv(R_REMOTES_NO_ERRORS_FROM_WARNINGS="true")
remotes::install_github("mfherman/gbatr")
```

## Using `gbat()`

### Specify return columns

By default, all 203 fields from GBAT functions 1A, 1E, and AP are returned when you call `gbat()`. You probably do not want all these columns in your geocoded data frame, so there are a few approaches to specifying which fields to return. 

#### Presets

First, there are a handful of "presets" you can specify that return the most frequently used GBAT fields. 

| Preset                   | GBAT Fields                      |
|:-------------------------|:---------------------------------|
| `"lot_centroid_xy"`      | F1A_Xcoordinate, F1A_Ycoordinate |
| `"lot_centroid_latlon"`  | F1A_Latitude, F1A_Longitude      |
| `"block_face_xy"`        | F1E_XCoordinate, F1E_YCoordinate |
| `"block_face_latlon"`    | F1E_Latitude, F1E_Longitude      |
| `"cd"`                   | F1E_CommunityDistrict            |
| `"nta"`                  | F1E_NTA, F1E_NTAName             |
| `"census_tract"`         | F1E_2010CensusTractGEOID         |
| `"census_block"`         | F1E_2010CensusBlockGEOID         |

You can use one of more of the presets by specifying them in the `geo_colnames` argument.

```{r}
gbat(
  za,
  address = "address",
  zip_boro = "borough",
  zip_boro_type = "boro",
  geo_colnames = c("cd", "nta", "census_tract")
  )
```

#### Column names

`gbat_fields` is a data set built into gbatr that contains all the possible GBAT fields you can request. If you want to specify one or more of these fields to be included in the `gbat()` output, use the name of that field from the `col_names` column in the `gbat_fields` data set.

```{r}
head(gbat_fields)
```

You can combine these column names with the presets discussed above.

```{r}
gbat(
  za,
  address = "address",
  zip_boro = "borough",
  zip_boro_type = "boro",
  geo_colnames = c("lot_centroid_latlon", "F1E_BBL")
  )
```

#### Geosupport functions 

A third option for choosing return columns is to use the `func` argument. You can read all about Geosupport Functions in the [Geosupport doucmentation](https://nycplanning.github.io/Geosupport-UPG/appendices/appendix01/#introduction).Function 1A and Function 1E return slightly different versions of similar information. For instance, you might have noticed that there are both 1A and 1E latitude and longitude fields. There is lots of detail in the Geosupport manual, but the important thing to know is that Function 1A returns tax lot- and building-level data while Function 1E returns block face-level data. In practice this means that when you geocode an address, the coordinates returned by F1A will generally be the centroid of the tax lot, while the coordinates returned by F1E will be a point on the street. For smaller, residenital buildings, this generally does not make a big difference, but if you are geocoding an address for a large housing development with multiple buildings, there may be larger differences between the points returned by 1A and 1E.

Here, for example, we get latitude and longitude for our pizza places from both 1A and 1E. Note the small differences between the latidues and longitudes returned.

```{r}
gbat(
  za,
  address = "address",
  zip_boro = "borough",
  zip_boro_type = "boro",
  geo_colnames = c("lot_centroid_latlon", "block_face_latlon")
  )
```

### Dealing with failure

### Converting to `sf` objects and mapping

```{r, eval=FALSE}
za_geo %>%
  sf::st_as_sf(
    coords = c("F1A_Longitude", "F1A_Latitude"),
    crs = 4326
    ) %>%
  mapview::mapview()
```





